---
title: "Project 1"
author: "Andy Lopez  \n Artem Plekhov  \n Shreya L Ramesh"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    toc_depth: 3

---
```{r setup, include = FALSE}
```

## Background
The World Health Organization has recently employed a new data science initiative, *CSIT-165*, that uses data science to characterize pandemic diseases. 
*CSIT-165* disseminates data driven analyses to global decision makers.

*CSIT-165* is a conglomerate comprised of two fabricated entities: *Global Health Union (GHU)* and *Private Diagnostic Laboratories (PDL)*. 
Your and your partner's role is to play a data scientist from one of these two entities.

## Data
> [2019 Novel Coronavirus COVID-19 (2019-nCoV) Data Repository by John Hopkins CSSE](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series)

Data for 2019 Novel Coronavirus is operated by the John Hopkins University Center for Systems Science and Engineering (JHU CSSE).
Data includes daily time series CSV summary tables, including confirmations, recoveries, and deaths. 
Country/region are countries/regions hat conform to World Health Organization (WHO).
Lat and Long refer to coordinates references for the user. 
Date fields are stored in MM/DD/YYYY format.

## Project Objectives

### Objective 1
```{r echo=TRUE}
# read data
confirmed <- read.csv("time_series_covid19_confirmed_global.csv", header=TRUE)
print("The highest confirmed cases in day 1 is as follows:")
print(confirmed[which.max(confirmed[,5]),1:5])

deaths <- read.csv("time_series_covid19_deaths_global.csv", header=TRUE)
print("The highest death cases in day 1 is as follows:")
print(deaths[which.max(deaths[,5]),1:5])

recovered <- read.csv("time_series_covid19_recovered_global.csv", header=TRUE)
print("The highest recovered cases in day 1 is as follows:")
print(recovered[which.max(recovered[,5]),1:5])

if (confirmed[which.max(confirmed[,5]),1:2]==deaths[which.max(deaths[,5]),1:2] && confirmed[which.max(confirmed[,5]),1:2]==recovered[which.max(recovered[,5]),1:2]) {
  print("Since this area had the most cases, deaths, and recoveries in the first day, it is likely the origin")
} else {
  print("It is currently not possible to describe the origin of the disease")
}

```

### Objective 2
```{r ob2}

```

### Objective 3
```{r ob3}

```

### Objective 4

#### Objective 4.1
```{r ob4.1}

```

#### Objective 4.2
```{r ob4.2}

```

